---
# date: "`r Sys.Date()`"
output: github_document
header-includes:
  - \usepackage{amsfonts,amsmath,amsthm,amsbsy,amssymb,dsfont,bm,mathtools,mathalfa}
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  tidy = TRUE, 
  tidy.opts = list(comment = FALSE),
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
library(gdim)
```

# `gdim`: Graph dimensionality

<!-- badges: start -->
[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
<!-- badges: end -->

`gdim` is an R package for estimating graph dimensionality `k` (e.g., the number of communities in a network). The key function `eigcv` uses the cross-validated eigenvalues to test the statistical significance of sample eigenvectors. The test statistics enjoy a simple central limit theorem. As such, `eigcv` provides p-values for individual hypothesized `k`.  


## Installation

You can install from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("RoheLab/gdim")
```

<!-- You can install the released version of epca from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("gdim") -->
<!-- ``` -->

<!-- or the development version from [GitHub](https://github.com/) with: -->

<!-- ``` r -->
<!-- # install.packages("devtools") -->
<!-- devtools::install_github("RoheLab/gdim") -->
<!-- ``` -->

## Example

`eigcv()` is the main function in `gdim`. The function is simple to use and easy to configurable. The single required parameter for the function is the maximum possible dimension, `k_max`. 

In the following example, we generate a random graph from the stochastic block model (SBM) with 1000 nodes and 5 blocks. 

```{r sbm}
library(fastRG)
B = matrix(0.1, 5, 5)
diag(B) = .3
sbm <- sbm(n = 1000, k = 5, B = B, expected_degree = 40)
A <- sample_sparse(sbm, poisson_edges = T, allow_self_loops = F)
```
Here, `A` is the adjacency matrix.

Now, we call the `eigcv` function with `k_max=10` to estimate graph dimensionality. 
```{r eigcv}
eigcv(A, k_max = 10)
```
In this example, `eigcv()` suggests to choose `k=5`. 

<!-- For more examples, please see the vignette:  -->
<!-- ```{r, eval=FALSE} -->
<!-- vignette("epca") -->
<!-- ``` -->


## Getting help

If you encounter a clear bug, please file an issue with a minimal reproducible example on [GitHub](https://github.com/RoheLab/gdim/issues). 


## Reference

Chen F, Roch S, Rohe K, & Yu S. "Estimating graph dimension with cross-validated eigenvalues." *In preparation*.

